\TESS Photometric Mapping of a Terrestrial Planet in the Habitable Zone: Detection of Clouds, Oceans, and Continents

### 2019-03-28 tex/ms.tex
better abstract

\usepackage{xcolor}
\usepackage{xspace}
\newcommand{\TESS}{\emph{TESS}\xspace}
\newcommand{\todo}[1]{\textcolor{red}{#1}}

@@

\begin{abstract}
\todo{[maybe change first sentences to focus on planet mapping \& give the game 
away less?]} The Transiting Exoplanet Survey Satellite (\TESS) mission is a 
targeted effort to detect planets smaller than Neptune around bright, nearby stars. 
While \TESS is already enjoying great success with the discovery of many new worlds, 
the strongest signal in its data is typically ignored, as it lurks in the background 
of every camera pixel. In this work, we extract this signal and demonstrate that 
it is consistent with a terrestrial planet with a rotation period of 1 day. 
Using a spherical harmonic-based reflection model developed as an extension of 
the \starry package, we are able to reconstruct the surface features of this rocky 
world. We recover a time-variable albedo map of the planet including persistent 
regions which we interpret as continental features and cloud banks. 
We argue that this planet represents the most promising detection of a habitable 
world to date, although the potential intelligence of any life on it is yet to 
be determined.
\end{abstract}

@@

Of the many challenges and opportunities facing the field of exoplanets in the future, the prospect of mapping the surface of an Earth-like planet is one of the most exciting -- and one of the most difficult. 
Present-day missions including the Transiting Exoplanet Survey Satellite \citep[\TESS; ][]{Ricker2015} aim to discover small terrestrial planets around bright, nearby stars. 
Once these targets are found, efforts to characterize them in detail and determine their habitability will begin in earnest. 
The discovery of surface features including continents and oceans will be a critical hint for the habitability of an alien world.

Such mapping can be done by constructing a phase map of the planet using precise photometry obtained over long periods of time. 
\todo{(literature background \& citations)}

@@

A new source of precise photometry is the Transiting Exoplanet Survey Satellite 
\citep[\TESS; ][]{Ricker2015}. 
Launched in April 2018, \TESS has... \todo{(summary of \TESS's capabilities, 
citations to planet discoveries etc)}

One intriguing feature of the \TESS photometric data products is a highly 
time-variable signal represented in the diffuse background across virtually 
all camera pixels. 
\todo{(talk abt Earthshine without naming it)} 
The signal has been widely attributed to a planet which we will call Sol d.

Previous attempts to map this planet from the exoplanet community have suggested 
the presence of localized surface features, but their resolving power has been 
severely limited by the duration of observations \citep{Cowan2009}. 
\TESS offers precise 2-minute cadence data spanning a wide range of illumination 
phases, enabling a level of spatial and temporal resolving power that is 
unprecedented in the exoplanetary phase mapping literature.

Moreover, recent advances in fast analytic computation of spherical harmonics 
have made the reconstruction of such detailed maps more feasible than ever. 
\todo{(STARRY)}

In this work, we build on this methodological foundation to perform 
reflected-light mapping of planetary surface features on Sol d. 
We give an overview of the \TESS data used in \S\ref{sec:data}. 
In \S\ref{sec:methods}, we describe the methods employed to infer a map 
from these data, including the adaptation of the \starry algorithm to reflection 
mapping, the modeling of spacecraft-related systematics, and the likelihood and 
priors used. 
We present results in \S\ref{sec:results} and make comparison of these 
findings to state-of-the-art terrestrial planet maps from the literature in 
\S\ref{sec:discussion}. 
Finally, we conclude in \S\ref{sec:conclusions} with a look at future 
prospects for phase-curve mapping of terrestrial exoplanets.

@@

The \TESS pipeline produces processed lightcurves for all short-cadence targets. 
Part of this processing is the determination of the localized background flux 
within each postage stamp, which is provided as an ancillary time series. 
As with all \TESS time series data products, the time is given in \TESS Julian 
Days (TJD), defined as $\mathrm{BJD} - 2457000$, and the data are delivered in 27-day 
parcels corresponding to the varying pointing sectors of \TESS. 
We use these background flux measurements from Sectors 1 and 2 as our primary 
data source in this work.

@@

We have used \TESS photometry to
produce a global map of Sol d, a terrestrial planet in the habitable
zone of a G2 main-sequence star. We adapted the \starry code to 
generate (and invert) light curves in reflected light, coupling it
to a variation of pixel level decorrelation to model the systematics
in the \TESS detector. We detect three large scale bright regions near
the equator of the planet, associated with persistent cloud structures
at longitudes of $-60^\circ$, $+30^\circ$, and $+90^\circ$, which
likely track continents. Pronounced dark features are identified
at longitudes of $\pm 180^\circ$, $-30^\circ$, and $+60^\circ$ with
significant latitudinal extent, likely tracking oceans. Because of the
vantage point of \TESS during Sectors 1 and 2, we are not sensitive to
high latitudes on the planet, although this may change with upcoming
sectors.

@@

\acknowledgements{\todo{(Artist acknowledgement)} The authors gratefully 
acknowledge Dan Foreman-Mackey, David W. Hogg, Ben Pope, \todo{and ...} 
for useful conversations. Crucial parts of this work were carried out at the 
TESS Data Workshop, hosted by Space Telescope Science Institute, and the 
Building Early Science with TESS Workshop, hosted by the University of Chicago.}
\facility{TESS}
\software{Astropy, Matplotlib, Numpy, \starry \citep{Luger2019}, 
TensorFlow \citep{Abadi2015}, \spiceypy \citep{Annex2019}}

>>>>

\usepackage{xcolor}
\newcommand{\todo}[1]{\textcolor{red}{#1}}

@@

\begin{abstract}
Multiple exoplanets have now been photometrically mapped using phase-modulated reflection or emission from their surfaces and atmospheres, but the small amplitudes of such signals have limited previous maps almost exclusively to hot gas giant planets. 
In this work, we uncover a signal using recently released data from the Transiting Exoplanet Survey Satellite (\TESS) which we show corresponds to time-variable reflection from a terrestrial 
planet with a rotation period of $0.9972696$ days. 
Using a spherical harmonic-based reflection model developed as an extension of 
the \starry package, we are able to reconstruct the surface features of this rocky 
world. We recover a time-variable albedo map of the planet including persistent 
regions which we interpret as oceans and cloud banks indicative of continental features. 
We argue that this planet represents the most promising detection of a habitable 
world to date, although the potential intelligence of any life on it is yet to 
be determined.
\end{abstract}

@@

Of the many challenges and opportunities facing the field of exoplanets in the future, the prospect of mapping the surface of a potentially habitable planet is one of the most exciting -- and one of the most difficult. 
Present-day missions including the Transiting Exoplanet Survey Satellite \citep[\TESS; ][]{Ricker2015} aim to discover small terrestrial planets around bright, nearby stars. 
Once these targets are found, much more detailed observational characterization must follow to evaluate their suitability for harboring life. 
The discovery of surface features including continents and oceans would be a critical hint towards the habitability of an alien world.

The most straightforward method of observing features on a terrestrial exoplanet would be to directly image the planet at sufficient resolution. 
However,

Surface mapping can also be done by observing rotational phase-dependent variations of the planetary signal in precise photometric time series. 
\todo{(summarize how it works)} 
This is a complex data analysis problem: strong degeneracies are involved when transforming a 1-dimensional time series into a 2-dimensional surface map. 
Interpretation is also challenging in the face of many potential explanations for a poorly resolved observed feature.

@@

%A new source of precise photometry is the Transiting Exoplanet Survey Satellite 
%\citep[\TESS; ][]{Ricker2015}. 
%Launched in April 2018, \TESS has... \todo{(summary of \TESS's capabilities, 
%citations to planet discoveries etc)}

%One intriguing feature of the \TESS photometric data products is a highly 
%time-variable signal represented in the diffuse background across virtually 
%all camera pixels. 
%\todo{(talk abt Earthshine without naming it)} 
%The signal has been widely attributed to a planet which we will call Sol d.

Previous attempts to map this planet from the exoplanet community have suggested 
the presence of localized surface features, but their resolving power has been 
severely limited by the duration of observations \citep{Cowan2009}. 
\TESS offers precise 2-minute cadence data spanning a wide range of illumination 
phases, enabling a level of spatial and temporal resolving power that is 
unprecedented in the history of exoplanet phase mapping.

Moreover, recent advances in the computational tractability of this data analysis problem 
have made the reconstruction of detailed maps more feasible than ever. 
\todo{(describe how STARRY works - fast analytic computation of spherical harmonics)}

In this work, we build on this methodological foundation to perform 
reflected-light mapping of planetary surface features on Sol d. 
We give an overview of the \TESS data used in \S\ref{sec:data}. 
In \S\ref{sec:methods}, we describe the methods employed to infer a map 
from these data, including the adaptation of the \starry algorithm to reflection 
mapping, the modeling of spacecraft-related systematics, and the likelihood and 
priors used. 
We present results in \S\ref{sec:results} and make a comparison of these 
findings to state-of-the-art terrestrial planet maps in 
\S\ref{sec:discussion}. 
Finally, we conclude in \S\ref{sec:conclusions}.

@@

The \TESS pipeline produces processed lightcurves for all short-cadence targets. 
Part of this processing is the determination of the localized background flux 
within each postage stamp, which is provided as an ancillary time series. 
As with all \TESS time series data products, the time is given in units of \TESS Julian 
Date (TJD), defined as $\mathrm{BJD} - 2457000$ \todo{(is this barycentric corrected?)}, and the data are delivered in 27-day 
parcels corresponding to the varying pointing sectors of \TESS. 
We use these background flux measurements from Sectors 1 and 2 as our primary 
data source in this work.

@@

We have used \TESS photometry to
produce a global map of Sol d, a terrestrial planet in the habitable
zone of a G2 main-sequence star. We adapted the \starry code to 
generate (and invert) light curves in reflected light, coupling it
to a variation of pixel level decorrelation to model the systematics
in the \TESS detector.

We detect three large scale bright regions near
the equator of the planet, associated with persistent cloud structures
at longitudes of $-60^\circ$, $+30^\circ$, and $+90^\circ$, which
likely track continents. Pronounced dark features are identified
at longitudes of $\pm 180^\circ$, $-30^\circ$, and $+60^\circ$ with
significant latitudinal extent, likely tracking oceans. Because of the
vantage point of \TESS during Sectors 1 and 2, we are not sensitive to
high latitudes on the planet, although this may change with upcoming
sectors.

@@

In conclusion, we have demonstrated that \TESS is an effective addition to NASA's fleet of weather satellites.

\acknowledgements{\todo{(Artist acknowledgement)} The authors gratefully 
acknowledge Jonathan Fraine, Dan Foreman-Mackey, David W. Hogg, Ben Montet, and Ben Pope 
for useful conversations. Crucial parts of this work were carried out at the 
TESS Data Workshop, hosted by Space Telescope Science Institute, and the 
Building Early Science with TESS Workshop, hosted by the University of Chicago.}
\facility{TESS}
\software{Astropy, Matplotlib, Numpy, \starry \citep{Luger2019}, 
TensorFlow \citep{Abadi2015}, \spiceypy \citep{Annex2019}}
